---
title: "Loops"
author: "Blake Tindol"
date: "6/14/2020"
output: html_document
---
## Loops using a vector
Create a random vector for manipulation
```{r}
# Loops Working with a vector
set.seed(123) # ensure it is reproducible
df <-data.frame(sample(15:35, 10, replace=FALSE))
colnames(df)[1] <- "c1" # make column one called c1
```

## Looping over the dataframe column
Print out every value
``` {r}
for(i in 1:length(df$c1)){
  print(i)
}
```

# Minus one from every value in column
```{r}
for(i in 1:length(df$c1) -1) {
  print(i)
}

for(i in 1:length(df$c1) +1) {
  print(i)
}


```

## Create two column dataframe
```{r}
df2 <-data.frame(sample(15:35, 10, replace=FALSE))                 
df3 <-data.frame(sample(15:35, 10, replace=FALSE))
df4 <- cbind(df2,df3)
colnames(df4)[1]<-"c1"
colnames(df4)[2]<-"c2"
```

## Multiplication of values
Make new column by multipling 2 columns 
```{r}
for(i in 1:length(df4)){
  df4$c3 <- df4$c1*df4$c2
}
head(df4)

```

## Nested Loops 
Print loop
```{r}
df2 <-data.frame(sample(15:35, 10, replace=FALSE))                 
df3 <-data.frame(sample(15:35, 10, replace=FALSE))
df4 <-data.frame(sample(15:35, 10, replace=FALSE))
df5<-data.frame(sample(15:35, 10, replace=FALSE))
df6 <-data.frame(sample(15:35, 10, replace=FALSE))
df <- cbind(df2,df3,df4,df5,df6)
colnames(df)[1]<-"c1"
colnames(df)[2]<-"c2"
colnames(df)[3]<-"c3"
colnames(df)[4]<-"c4"
colnames(df)[5]<-"c5"


for(row in 1:nrow(df)) {
  for(col in 1:ncol(df)) {
    print(df[row, col])
  }
}

# Just to understand the make up of the dataframe for further use
print(df)
```



## More Nested Loops
```{r}
# Create a nested loop
for(row in 1:nrow(df)) {
  for(col in 1:ncol(df)) {
    print(paste(colnames(df)[col], "and", rownames(df)[row],
                "have a value of", df[row,col]))
  }
}
```
 
# Shifting Positions
```{r}
# Create a nested loop
for(row in 1:nrow(df) - 1) { # This shifts the whole data frame up one row that is why there is no value to print
  for(col in 1:ncol(df)) {
    print(paste(colnames(df)[col], "and", rownames(df)[row],
                "have a correlation of", df[row,col]))
  }
}
```

# Shifting Continued
``` {r}
# Create a nested loop
for(row in 1:nrow(df)) { 
  for(col in 1:ncol(df)+1) {# This shifts the whole data frame over to the right one that is why there is no far left column value to print - 1 will shift dataframe over the right left 1 making far right column empty
    print(paste(colnames(df)[col], "and", rownames(df)[row],
                "have a correlation of", df[row,col]))
  }
}
```


## Simulation of random sampling 
```{r}
df <- rep(NA, 10000) # make vector of 10000 empty NA cells
for(i in 1:10000){
    a <- rpois(50, lambda=20) # possion distribution lambda is center value (median)
    df[i] <-mean(a) # take the mean of the 50 differnt values 
}
hist(df) # make histograme
```